---
title: "Project 1"
author: "Tanja Venborg"
date: "4 August 2016"
output: html_document
---

# Loading the data

First the activity data will be loaded

```{r load data, echo = TRUE}
data <- read.csv("/Users/Thansen/Desktop/R/ReproducibleResearch/activity.csv")
data <- na.omit(data)
```


# Processing data

Here I find the sum of the steps taken each day. I start by looking at the data. Then I create a new datatable, where the steps are summed up by date, then I add the columnnames. Finally I print the new steps table, where its now clear how many steps where taken each of the dates where ther is a step. 

```{r}
head(data)
library(data.table)
stepstabel <-  as.data.table(data)[, by = . (date), sum(steps)]
colnames(stepstabel) <- c("Date","steps")
stepstabel$Date <- format(as.Date(stepstabel$Date), "%Y/%m/%d")
print(stepstabel)
```

# Creating a histogram

```{r, echo = TRUE}
steps1 <- hist(stepstabel$steps, breaks =20, main = "Total steps per day", xlab ="Steps", ylab =  "Frequency", col ="green")
```

# Plotting the daily activity pattern

To create the plot, i first stort the data in a sum of step by the intervals. To get the line plot, I type = "l". If this isn't specified, the plot will be a scatterplot.
```{r}
interval <- as.data.table(data)[,sum(steps), by = .(interval)] 
colnames(interval) <- c("interval", "steps")
plot(interval, type = "l")
```

Finding the interval with the maximum numbers of steps is pretty straight forward.
``` {r}
maxstep <- sapply (interval, max)
interval[steps %in% maxstep,]
                            
```

This means that interval 835 has the largest amount of steps.

# Imputing missing values

Since I removed the NA's from the first dataset, I reload it and in a new data name. 
Calculating the total amount of missing values from the original dataset
```{r}
orgdata <- read.csv("/Users/Thansen/Desktop/R/ReproducibleResearch/activity.csv")
sum(is.na(orgdata))
nrow(orgdata)
sum(is.na(orgdata))/nrow(orgdata)
```

There was 2304 missing values. That seems like quite a lot, actually it is 13%.  

To fill in the missing values, I take the mean of the filled-in days.

```{r}
orgdata$steps[is.na(orgdata$steps)] <-  mean(data$steps, na.rm = FALSE)
head(orgdata)
sum(is.na(orgdata))
```

There is no longer any missing values in the data.

```{r}
filled_in_data <-  as.data.table(orgdata)[, by = . (date), sum(steps)]
colnames(filled_in_data) <- c("interval", "steps")
steps2 <- hist(filled_in_data$steps, breaks = 20, xlab = "Steps", main = "The total amount of steps with filled in data", col ="red")
```

Combining the two histograms to see the difference

```{r}
set.seed(42)
plot(steps1, col = rgb(0,0,1, 1/4), xlim = c(0,24000))
plot(steps2, col = rgb(1,0,0,1/4), xlim = c(0,24000), add=T)
```

# Weekdays vs weekends
First I'm adding and extra row where the weekend, weekday distinguishion is made.

```{r}
orgdata$realdate <-  as.Date(orgdata$date, format = "%Y-%m-%d")
orgdata$weekday <- weekdays(orgdata$realdate)

orgdata$daytype <- ifelse(orgdata$weekday == "Saturday" | orgdata$weekday == "Sunday", "weekend", "weekday")
head(orgdata, n = 10) 
```

Plotting the weekdays vs. weekends

```{r}
sumdata_interval <- aggregate(steps~interval+daytype,data = orgdata, FUN = mean, na.action = na.omit)
head(sumdata_interval)
library(lattice)
xyplot(steps ~interval | daytype, data= sumdata_interval, layout =c(1,2), type = "l")
```